---
title: "Rexample"
author: "Lingxiao Zhou"
date: "2024-09-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning = FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60))
```



# Bollywood Box Office Data

* Movie Budgets (X)
* Box Office Grosses (Y)

```{r fig.height=3, fig.width=7}
library(car) # for significance tests
bbo <- read.csv("https://raw.githubusercontent.com/lingxiaozhou/STA4210Rmaterial/main/data/bollywood_boxoffice.csv",
header=T)
attach(bbo)
names(bbo)
bbo.reg1 <- lm(Gross ~ Budget)
summary(bbo.reg1)
e1 <- residuals(bbo.reg1)
yhat1 <- predict(bbo.reg1)

par(mfrow=c(1,2))

# Residual vs. fitted
plot(yhat1,e1,main="Residuals vs Fitted Values", xlab="Fitted Values",
ylab="Residuals")
abline(h=0)

# QQ plot
qqnorm(e1)
qqline(e1)


shapiro.test(e1) # Shapiro-Wilk
ncvTest(bbo.reg1) # Breusch-Pagan
```

* Diagnostic plots and the significance tests indicate that there are violations to normality and constant variance assumption.


## Log transformation of Y

```{r fig.height=3, fig.width=9}
bbo.reg2 <- lm(log(Gross) ~ Budget)
summary(bbo.reg2)
e2 <- residuals(bbo.reg2)
yhat2 <- predict(bbo.reg2)

par(mfrow=c(1,3))

# Scatter plot
plot(Budget, log(Gross))
abline(bbo.reg2)

# Residual vs. fitted
plot(yhat2,e2,main="Residuals vs Fitted Values", xlab="Fitted Values",
ylab="Residuals")
abline(h=0)

# QQ plot
qqnorm(e2); qqline(e2)

shapiro.test(e2) # Shapiro-Wilk
ncvTest(bbo.reg2) # Breusch-Pagan
```

* Based on the Shapiro-Wilk test (Normality) and the Breusch-Pagan test (Constant Variance), the new model appears to be better. However, the linearity assumption is violated based on the residuals vs. fitted plot.


## Log transformation of x and Y

```{r fig.height=3, fig.width=9}
bbo.reg3 <- lm(log(Gross) ~ log(Budget))
summary(bbo.reg3)
e3 <- residuals(bbo.reg3)
yhat3 <- predict(bbo.reg3)
par(mfrow=c(1,3))

# Scatter plot
plot(log(Budget), log(Gross))
abline(bbo.reg3)

# Residual vs. fitted
plot(yhat3,e3,main="Residuals vs Fitted Values", xlab="Fitted Values",
ylab="Residuals")
abline(h=0)

# QQ plot
qqnorm(e3); qqline(e3)

shapiro.test(e3) # Shapiro-Wilk
ncvTest(bbo.reg3) # Breusch-Pagan
```




## Box-Cox transformation

* $W = \begin{cases}\frac{Y^\lambda-1}{\lambda} & \lambda\neq 0\\\ln(Y) & \lambda=0\end{cases}$

```{r fig.height=3, fig.width=3}
library(MASS) # for boxcox
bbo.reg4 <- lm(Gross ~ Budget)
bc <- boxcox(bbo.reg4,plotit=T)
lambda <- bc$x[which.max(bc$y)]
lambda


#fit new linear regression model using the Box-Cox transformation
bbo.reg.bc <- lm(((Gross^lambda-1)/lambda) ~ Budget)
```

* The optimal $\lambda$ is 0.22
* The procedure chooses a “quarter root” transformation for Y. We will not pursue that here, as we have seen that log transformations of Y and X work quite well.



# Lowess

* Nonparametric method of obtaining a smooth plot of the regression relation between Y and X
* Fits regression in small neighborhoods around points along the regression line on the X axis
* Weights observations closer to the specific point higher than more distant points
* Re-weights after fitting, putting lower weights on larger residuals (in absolute value)
* Obtains fitted value for each point after “final” regression is fit
* Model is plotted along with linear fit, and confidence bands, linear fit is good if lowess lies within bands


```{r fig.height=4, fig.width=9}
par(mfrow=c(1,2))


# find relationship between lot size (X) and work hours (Y) 
toluca=read.table("http://www.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/KutnerData/Chapter%20%201%20Data%20Sets/CH01TA01.txt",
                  col.names = c("lotsize", "workhrs"))
plot(toluca$lotsize,toluca$workhrs, xlab = "lotsize", ylab = "workhrs")
toluca.reg=lm(workhrs~lotsize,data=toluca)

abline(toluca.reg,col="darkgreen")


x_seq=seq(20,120,by=1)
fitl=loess(workhrs~lotsize,span=0.5,data=toluca) # span controls the size of the neighborhood
predl=predict(fitl,x_seq,se=TRUE)   # Get predicted y values for x_seq based on lowess model

plot(toluca)
lines(x_seq,predl$fit,lty=1,col="darkred")
lines(x_seq,predl$fit-1.96*predl$se.fit, lty=2, col="blue", lwd=1)
lines(x_seq,predl$fit+1.96*predl$se.fit, lty=2, col="blue", lwd=1)

polygon(c(x_seq,rev(x_seq)),c(predl$fit+1.96*predl$se.fit,rev(predl$fit-1.96*predl$se.fit)),col="#00009933",border="NA")

abline(toluca.reg,col="darkgreen")
legend("bottomright",legend=c("Loess","95% CB","SLR"),
       col=c("darkred","blue","darkgreen"),lty=c(1,2,1),bty="n")
```



