---
title: "Rexample5"
author: "Lingxiao Zhou"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning = FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60))
```

```{r fig.height=3.2, fig.width=5}
# find relationship between lot size (X) and work hours (Y) 
toluca <- read.table("http://www.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/KutnerData/Chapter%20%201%20Data%20Sets/CH01TA01.txt",
                  col.names = c("lotsize", "workhrs"))
plot(toluca$lotsize,toluca$workhrs,xlab = "Lot size", ylab = "Work hours")
toluca.reg <- lm(workhrs~lotsize,data=toluca)
summary(toluca.reg)
abline(toluca.reg)
```

```{r}
e <- toluca.reg$residuals # get the raw residuals
```


```{r echo=FALSE}
### Checking model assumptions graphically
par(mfrow=c(2,2))

# Normality
hist(e, main = "Histogram of residuals", xlab = "residual")
qqnorm(e,datax=TRUE)
qqline(e,datax=TRUE)

# Independence
plot(e,type="o",pch=22,xlab="Order",main="Independence", ylab = "residuals")
abline(h=0)

# Homogeneity of variance/Model Fit
plot(e~fitted.values(toluca.reg),main="Homogeneity/Fit",xlab = "fitted values", ylab = "residuals")
abline(h=0)
```




# Normality

## Shapiro test

* $H_0:$ The residuals are drawn from normal distribution
* Since P-value > 0.05, we fail to reject the normal assumption at the significance level 0.05


```{r}
shapiro.test(e)
```


# Independence

## runs test

* $H_0:$ Independence
* We fail to reject the null hypothesis since p-value is large. 
* According to the plot, there are 11 runs out of maximum 25


```{r fig.height=3.5, fig.width=6}
par(mfrow=c(1,1))
library(lawstat) # may need to install package
runs.test(e,plot.it=TRUE)
```

## Durbin-Watson

* Assume $\epsilon_i = \rho\epsilon_{i-1}+u_i$ where $u_i\stackrel{iid}{\sim}N(0,\sigma^2).$
* $H_0:\rho = 0$
* For two-sided test, $H_a:\rho\neq0$
* Notice that `durbinWatsonTest()` and `dwtest()` use different method to calculate p-value, so you may observe that they did not give consistent results.
* Here we fail to reject the $H_0$.

```{r}
set.seed(1234)

library(car) # for durbinWatsonTest
durbinWatsonTest(toluca.reg,alternative="two.sided")


library(lmtest) # for dwtest
dwtest(toluca.reg,alternative="two.sided")
```


# Homogeneity of Variance

## Levene's test

* Need categorical x
* $H_0:\sigma^2_1 = \sigma_2^2=\dots=\sigma_t^2$  
* Split the data into two groups depending on whether lot size is greater
than 75 or not (half-way point).
* With a p-value greater than 0.05 we fail to reject the null.

```{r}
ind <- I(toluca$lotsize>75) # convert lotsize to categorical variable
temp <- cbind(toluca$lotsize,e,ind);temp  # print X, residual and converted X

leveneTest(temp[,2],ind)
```

## Breusch-Pagan/Cook-Weisberg test

* $H_0:$ The residuals are distributed with equal variance
* Fail to reject the null since the p-value = 0.36491 > 0.05

```{r}
ncvTest(toluca.reg) # car library
```

# Linearity of regression

## Lack of fit

* $H_0: E(Y_i) = \beta_0+\beta_1X_i$
* $H_a: E(Y_i) \neq \beta_0+\beta_1X_i$
* P-value = 0.6893. We fail to reject the null hypothesis and there is no violation for the linearity assumption based on this test

```{r}
# check t<<n
# If t is approximately equal to n, then the test is not applicable
(t <- length(unique(toluca$lotsize)))
(n <- nrow(toluca))

Reduced <- toluca.reg # reduced model: SLR model
Full <- lm(workhrs~0+as.factor(lotsize),data=toluca) # full model: use group mean
summary(Full)
anova(Reduced, Full) # get lack-of-fit test
```

