---
title: "Rexample1"
author: "Lingxiao Zhou"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=60))
```


# Copier example

Let X be the number of copiers serviced and Y be the time spent (in
minutes) by the technician for a known manufacturer.


```{r}
copier <- read.csv("https://raw.githubusercontent.com/lingxiaozhou/STA4210Rmaterial/main/data/copiers.csv",skip=1,header=TRUE)

head(copier)  # print the first six rows of the dataframe
nrow(copier)  # check the sample size
```








```{r fig.height=3.5, fig.width=7}
# Scatterplot
plot(copier$Copiers,copier$Time,pch=16,xlab="Quantity",ylab="Time (in minutes)",main="Scatterplot")


# Fit the model
reg <- lm(Time~Copiers,data=copier)
summary(reg)  # get the summary of the model
abline(reg)   # add the fitted line to the scatterplot
```



The estimated equation is
$$\hat Y = -0.5802 + 15.0352X,\ \mathrm{for}\  X \in \mathrm{approximately}\   [0, 10]$$

Recall that for simple linear regression model, we have
$$Y_i = \beta_0+\beta_1X_i+\epsilon_i,\hspace{1cm}\epsilon_i\sim N(0,\sigma^2),$$ which implies
$$E[Y_i] = \beta_0+\beta_1X_i$$

* We note that the slope $b_1 = 15.0352$ implies that for each unit
increase in copier quantity, the mean service time increases by 15.0352
minutes (for quantity values between 1 and 10).
* The estimated $\sigma$ is s = 8.914 from the output.
* If we wish to estimate the expected time needed for a service call for 5 copiers
that would be $-0.5802 + 15.0352(5) = 74.5958 \  \mathrm{minutes}.$

```{r}
predict(reg,newdata=data.frame(Copiers=5))
```

```{r fig.height=3, fig.width=5}
# print the residuals
cat("first 5 residuals:", reg$residuals[1:5],"\n")
plot(reg$residuals,main = "",ylab = "residual")
```

* The first residual implies that the actual observed time was `r abs(round(reg$residuals[1],4))` minutes smaller than the model estimates.
* Residual magnitude in minute may not be easy to interpret whether large or small in the context of the problem.
* standardized residual can be used to compare the residuals from different models: $\frac{e_i}{\mathrm{sd}(e_i)}$
    - standardized residuals are centered and scaled to make it easier to interpret residuals from different models.
    - An observation with a standardized residual that is larger than 3 (in absolute value) is considered big.

```{r}
cat("first 5 standardized residuals:",rstandard(reg)[1:5])
```



```{r}
copier$Time <- copier$Time/60               # convert the unit of time to hours
```


```{r}
# refit the model
reg2 <- lm(Time~Copiers,data=copier)   

# print the residuals
cat("first 5 residuals:", reg2$residuals[1:5],"\n")
cat("first 5 standardized residuals:",rstandard(reg2)[1:5])
```



