---
title: "Rexample"
author: "Lingxiao Zhou"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning = FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60))
```

# cruise dataset

## load data

```{r}
cruise <- read.fwf("https://raw.githubusercontent.com/lingxiaozhou/STA4210Rmaterial/main/data/cruise_ship.dat.txt", width=c(20,20,rep(8,7)),
                   col.names=c("ship", "cline", "age", "tonnage", "passengers", "length", "cabins", "passdens", "crew"))

head(cruise)


```

## Fit Full model

```{r}
fit0 <- lm(crew ~ age + tonnage + passengers + length + cabins + passdens,data=cruise)
summary(fit0)
AIC(fit0)
anova(fit0)

```
* Fit the linear regression model with all predictors
* According to the summary results and anova table, some predictors are not significant

## Perform all possible regressions (aka all subset regressions)


```{r}

# install.packages("leaps")
library(leaps)
allcruise <- regsubsets(crew ~ age + tonnage + passengers + length + cabins + passdens,nbest=4,data=cruise)
aprout <- summary(allcruise)
with(aprout,round(cbind(which,rsq,adjr2,cp,bic),3))     ## Prints "readable" results
plot(allcruise,scale="adjr2")
```


* The first column gives the number of predictors (p-1), not the number of parameters (p).
* The first row corresponds to the model with an intercept and cabins
* nbest = 4 tells the program to print out the best 4 models for each possible number of parameters. 
* The `with(aprout,round(cbind(which,rsq,adjr2,cp,bic),3))` command prints out the output in readable form, with 3 decimal places.
* Based on adjusted $R^2$, the best model is the model with tonnage, passengers, length and cabins



```{r}
fit.new <- update(fit0,.~.-age-passdens)
AIC(fit.new)

# install.packages("olsrr")
library(olsrr)

press_value <- ols_press(fit.new)
print(press_value)

```




