---
title: "Rexample3"
author: "Lingxiao Zhou"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=60))
```



# Example for ANOVA (toluca)

Data on lot size ($X$) and work hours ($Y$) was obtained from 25 recent runs of a manufacturing process.

```{r fig.height=3.5, fig.width=5}
# find relationship between lot size (X) and work hours (Y) 
toluca=read.table("http://www.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/KutnerData/Chapter%20%201%20Data%20Sets/CH01TA01.txt",
                  col.names = c("lotsize", "workhrs"))
plot(toluca$lotsize,toluca$workhrs, xlab = "lot size",ylab = "work hours")
toluca.reg=lm(workhrs~lotsize,data=toluca)
abline(toluca.reg)
```

```{r}
anova(toluca.reg)
summary(toluca.reg)
```

* Note note that the F-test and t-test for $\beta_1$ are equivalent and the p-values are the same.

## Perform the general linear test

**Reduce Model:** $Y_i = \beta_0$  
**Full Model:** $Y_i = \beta_0+\beta_1x_i$  
$H_0:$ Full model is not better than the reduced model (equivalently $\beta_1 = 0$)  
$H_a:$ Full model is better than the reduced model (equivalently $\beta_1\neq 0$)

```{r}
Red <- lm(workhrs~1,data=toluca)
Full <- toluca.reg

anova(Red,Full)
```

* P-value is $4.449\times10^{-10}<0.05.$ We reject null hypothesis and conclude that the full model is significantly better than the reduced model.

