---
title: "Rexample"
author: "Lingxiao Zhou"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning = FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60))
```

# EX (Drug)

* Response variable Y
* Dosage level (d)
* Drug product with levels: A, B, C
    - reference level: A
    - Dummy variables: P~B~, P~c~

## Create the dataset

```{r}
Dose=rep(c(0.2,0.4,0.8,1.6),each=3)
Product=rep(c("A","B","C"),4)
y=c(2.0,1.8,1.3,4.3,4.1,2.0,6.5,4.9,2.8,8.9,5.7,3.4)
ds=data.frame(Dose=(Dose),Product=factor(Product),Response=y)
xtabs(Response~Dose+Product,data=ds)
ds
```

## Fit multiple regression with interaction

$Y = \beta_0+\beta_1d+\beta_2P_B+\beta_3P_C+\beta_4dP_B+\beta_5dP_C+\epsilon$

```{r fig.height=3.5, fig.width=6}
modelfull=lm(Response~Dose*Product,data=ds)
summary(modelfull)
library(car)
scatterplot(Response~Dose|Product,smooth=FALSE,data=ds)
```

Notice that a log transformation on Dose may help fit a more linear relationship

```{r fig.height=3.5, fig.width=6}
ds=data.frame(logDose=log(Dose),Product=factor(Product),Response=y)
# Full model
modelfull=lm(Response~logDose*Product,data=ds)
summary(modelfull)
scatterplot(Response~logDose|Product,smooth=FALSE,data=ds)
```

## CI for coefficient

Create at least 90\% simultaneous CI for $\beta_4$, $\beta_5$ and $\beta_4-\beta_5$ using **Bonferroni method**

```{r}
coefficients(modelfull)
```

* $b_4 = -1.50$
* $b_5 = -2.28$
* $b_4-b_5 = -3.78$
* Critical value $t_{1-0.1/(2\cdot 3),n-p} = t_{1-0.1/6,6}$

```{r}
vmat=vcov(modelfull)
round(vmat,3)
```

* $S_{b_4-b_5} = \sqrt{S^2_{b_4}+S^2_{b_5}-2S_{b_4b_5}}$
* $S^2_{b_4}$ is the (5,5)th element in the matrix
* $S^2_{b_5}$ is the (6,6)th element in the matrix
* $S_{b_4b_5}$ is the (5,6)th element in the matrix


### CI for beta4

```{r}
coefficients(modelfull)[5]+c(-1,1)*qt(1-0.10/6,6)*sqrt(vmat[5,5])
```

### CI for beta5

```{r}
coefficients(modelfull)[6]+c(-1,1)*qt(1-0.10/6,6)*sqrt(vmat[6,6])
```


### CI for beta4-beta5

```{r}
coefficients(modelfull)[5]-coefficients(modelfull)[6]+
  c(-1,1)*qt(1-0.10/6,6)*sqrt(vmat[5,5]+vmat[6,6]-2*vmat[5,6])
```

